# Basic Vector practice
Creating 2 vectors of blood pressure, before and after treatment. Labeling by patient. Finding a difference between post treatment bp and pre treatment bp and its mean. Identifying which have positive change and their mean.
```{r Task 1}
pre<-c(130,128,116,124,133,134,118,126,114,127,141,138,128,140,137,131,120,128,139,135)
post<-c(114,98,113,99,107,116,113,111,119,117,101,119,130,122,106,106,124,102,117,113)

names(pre)<-c(paste("Subject", 1:20, sep="_"))
names(post)<-c(paste("Subject", 1:20, sep="_"))

change<-c(pre-post)

mean(change)

which(change>0)

positive<-which(change>0)

mean(positive)

```
# Data Frame Practice
Creating data frame of patient, pre, post, and change bp. Adding a new column that computes true if post bp is less than 120.
```{r Task 2}
bp_df<-data.frame("pre_bp"=pre,"post_bp"=post,"diff_bp"=change)

bp_df[bp_df$diff_bp < 0, ]

bp_df$low_bp<-c(bp_df$post_bp < 120)

knitr::kable(bp_df)

```
# List Practice
Creating new data set that is for a control group. Making a list with the treatment and placebo data frames. Accessing the first element 3 different ways. Then accessing the pre bp vector for the placebo group.
```{r Task 3}
placebo_pre<-c(138,135,147,117,152,134,114,121,131,130)
placebo_post<-c(105,136,123,130,134,143,135,139,120,124)
placebo_diff<-placebo_post-placebo_pre
names(placebo_pre)<-c(paste("Subject", 1:10, sep="_"))
names(placebo_post)<-c(paste("Subject", 1:10, sep="_"))

placebo_df<-data.frame("pre_bp"=placebo_pre, "post_bp"=placebo_post, "diff_bp"=placebo_diff)

placebo_df$low_bp<-c(placebo_df$post_bp < 120)

group<-list(treatment=bp_df,placebo=placebo_df)

group[1]
group[[1]]
group["treatment"]

group$placebo$pre_bp

```
# Control Flow Practice
Creating a column that categorizes post bp by optimal (≤ 120), borderline (120 < bp ≤ 130), and high (> 130) for both treatment and placebo.
```{r Task 4}
bp_df$status<-character(20)
placebo_df$status<-character(10)

for(i in 1:length(bp_df$post_bp)){
  if (bp_df$post_bp[i] >= 130){
  bp_df$status="high" 
    } else if (bp_df$post_bp[i] > 120) {
      bp_df$status="borderline" 
        } else if (bp_df$post_bp[i] <= 120) {
          bp_df$status="optimal"
        } else {
          bp_df$status="error"
        }
}

for(i in 1:length(placebo_df$post_bp)){
  if (placebo_df$post_bp[i] >= 130){
  placebo_df$status="high" 
    } else if (placebo_df$post_bp[i] > 120) {
      placebo_df$status="borderline" 
        } else if (placebo_df$post_bp[i] <= 120) {
          placebo_df$status="optimal"
        } else {
          placebo_df$status="error"
        }
}
```

# Functions
Writing a function to calculate statistics on pre, post, and difference bp for both treatment and placebo.
```{r Task 5}
stats<-function(bp_df, placebo_df, stat="mean"){
  my_fun<-get(stat)
  label<-c("treatment_pre", "treatment_post", "treatment_diff", "placebo_pre", "placebo_post", "placebo_diff")
  label_stat<-c(paste(label, stat, sep="_"))
  values<-c(my_fun(bp_df$pre_bp), my_fun(bp_df$post_bp), my_fun(bp_df$diff_bp), my_fun(placebo_df$pre_bp), my_fun(placebo_df$post_bp), my_fun(placebo_df$diff_bp))
  names(values)<-label_stat
  return(values)
}

stats(bp_df,placebo_df)
stats(bp_df,placebo_df, stat="var")
stats(bp_df,placebo_df, stat="sd")
stats(bp_df,placebo_df, stat="min")
stats(bp_df,placebo_df, stat="max")
```